---
title: "Barrett Viator DSE6111 - Preliminary Results (Resubmission)"
author: "Barrett Viator"
date: "2024-12-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Client Attrition Preliminary Results
When assessing the results from the preliminary logistic regression model and from KNN models against the 0.841 accuracy level of the null model, it is apparent
that further models will need to be explored to assess the risk of attrition based on past customer behaviors. The Logistic Regression model is performing well with an accuracy score of 0.9093 and and AUC of 0.9276. The The KNN AUC value of 0.8585 suggests that the KNN model is performing well with the classification process of Attrition Flags which also pairs well with the accuracy value of 0.8735 for the confusion matrix. 

Further processing of the data with subset selection and shrinkage methods applied has the advantage of providing ABC Corporation with better model accuracy as well
as the ability to narrow down which features are most greatly impacting the predictors. Applying boosting, bagging, and Random Forest Methods as well as exploring XG-boost will also enhance the ability to improve upon the preliminary models to move forward with final recommendations more precisely and soundly.

## Preparations, Data Cleaning, and Splitting

```{r, include = FALSE}
library(tidymodels)
library(parsnip)
library(kknn)
library(class)
library(yardstick)
library(dplyr)
library(ggplot2)
library(pROC)
library(ROCR)
library(ROSE)
library(vip)
library(caret)
library(e1071)
library(naivebayes)
library(MASS)
library(discrim)
library(pROC)
library(kknn)
library(ranger)

```

```{r, include= FALSE}
setwd("C:\\Program Files\\Merrimack\\DSE6111\\Project")
customer_data <- read.csv("C:\\Program Files\\Merrimack\\DSE6111\\Project\\customer_data.csv")
customer_data_na_removed <- na.omit(customer_data)
customer_data_clientnum_removed <- subset(customer_data_na_removed, select = -CLIENTNUM)

customer_data_split <- initial_split(customer_data_clientnum_removed, prop = 0.75)

training_set <- training(customer_data_split)
test_set  <- testing(customer_data_split)
training_set2 <- training(customer_data_split)
test_set2  <- testing(customer_data_split)
```

```{r, include= FALSE}
train_index <- createDataPartition(training_set$Attrition_Flag, p = 0.75, list = FALSE)

train_data <- training_set[train_index, ]
test_data <- training_set[-train_index, ]
train_data2 <- training_set[train_index, ]
test_data2 <- training_set[-train_index, ]
train_data3 <- training_set[train_index, ]
test_data3 <- training_set[-train_index, ]
train_data4 <- training_set[train_index, ]
test_data4 <- training_set[-train_index, ]
train_data5<- training_set[train_index, ]
test_data5 <- training_set[-train_index, ]
```

```{r}
train_data$Attrition_Flag <- as.factor(train_data$Attrition_Flag)

# Balance the data using oversampling (default method of ovun.sample)
balanced_data <- ovun.sample(Attrition_Flag ~ ., data = train_data, method = "over", N = 5697)$data

```

### Null Model

```{r}
null_model <- balanced_data %>%
  count(Attrition_Flag) %>%
  top_n(1, n) %>% 
  pull(Attrition_Flag)

null_predictions <- tibble(Attrition_Flag = rep(null_model, nrow(test_data)))


null_model_accuracy <- mean(null_predictions$Attrition_Flag == test_data$Attrition_Flag)
cat("Test Accuracy for the Null Model: ", null_model_accuracy, "\n")

```

### CV Sets and Recipes

```{r}
cv_set <- vfold_cv(train_data2, strata = "Attrition_Flag", v = 10)

customer_attrition_recipe <- recipe(Attrition_Flag ~ .,data = train_data2) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_center(all_predictors()) %>%
  step_scale(all_predictors())

```

### Logistic Regression 

```{r}
train_data$Attrition_Flag <- as.character(train_data$Attrition_Flag)
prepared_recipe <- prep(customer_attrition_recipe, training = train_data2)

processed_train_data <- bake(prepared_recipe, new_data = train_data2) 

customer_log_reg_mod <- 
  logistic_reg(penalty = tune(), mixture = 1) %>% 
  set_engine("glmnet")

customer_log_reg_workflow <- 
  workflow() %>% 
  add_recipe(customer_attrition_recipe) %>%
  add_model(customer_log_reg_mod)

customer_log_reg_grid <- tibble(penalty = 10^seq(-4, -1, length.out = 30))

log_reg_tuning_results <- tune_grid(
  object = customer_log_reg_workflow,
  resamples = cv_set,
  grid = customer_log_reg_grid,
  metrics = metric_set(roc_auc)
)

```

```{r} 

top_log_models <- log_reg_tuning_results %>% 
  show_best(metric = "roc_auc", n = 15) %>% 
  arrange(penalty)

top_log_models


```

```{r}
log_result_metrics <- collect_metrics(log_reg_tuning_results)

log_roc_auc_results <- log_result_metrics %>% 
  filter(.metric == "roc_auc")

print(log_roc_auc_results)

best_log_auc_model <- log_roc_auc_results %>% 
  arrange(desc(mean))

head(best_log_auc_model)

```

```{r}
train_data2$Attrition_Flag <- as.factor(train_data2$Attrition_Flag)

log_reg_spec <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")


log_reg_fit <- log_reg_spec %>%
  fit(Attrition_Flag ~ ., data = train_data2)

log_reg_pred_prob <- predict(log_reg_fit, test_data, type = "raw")
log_reg_pred_class <- ifelse(log_reg_pred_prob > 0.5, 1, 0)

log_reg_conf_matrix <- table(Predicted = log_reg_pred_class, Actual = test_data$Attrition_Flag)

print(log_reg_conf_matrix)

log_reg_accuracy <- sum(diag(log_reg_conf_matrix)) / sum(log_reg_conf_matrix)
print(paste("Accuracy: ", round(log_reg_accuracy, 3)))

```

```{r}

log_roc_curve <- roc(test_data$Attrition_Flag, log_reg_pred_prob)

log_auc <- auc(log_roc_curve)

print(paste("AUC:", log_auc))

```

```{r}
log_reg_roc_plot <- 
  log_reg_tuning_results %>% 
  collect_metrics() %>% 
  filter(.metric == "roc_auc") %>%  
  ggplot(aes(x = penalty, y = mean, color = penalty)) +  #
  geom_point() + 
  geom_line() +   
  ylab("ROC AUC") +  
  scale_x_log10(labels = label_number()) +  
  scale_color_continuous(type = "viridis", name = "Penalty") +  
  ggtitle("Attrition ROC AUC vs Penalty for Logistic Regression") + 
  theme_minimal() +  
  theme(plot.title = element_text(hjust = 0.5)) 

log_reg_roc_plot  

```

```{r}
train_data2$Attrition_Flag <- as.character(train_data2$Attrition_Flag)
log_best_model <- log_reg_tuning_results %>% 
  select_best(metric = "roc_auc")

log_final_model <- customer_log_reg_workflow %>%
  finalize_workflow(log_best_model) %>% 
  fit(data = train_data2) 


log_final_fit <- log_final_model %>%
  extract_fit_parsnip()


coefficients <- tidy(log_final_fit)


ggplot(coefficients, aes(x = reorder(term, estimate), y = estimate)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = "Feature Importance (Logistic Regression)",
       x = "Feature",
       y = "Coefficient Estimate") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8)) 


```

### KNN

```{r}
train_data$Attrition_Flag <- as.factor(train_data$Attrition_Flag)

customer_attrition_recipe4 <- recipe(Attrition_Flag ~ 
                                      Customer_Age + 
                                      Gender + 
                                      Dependent_count + 
                                      Education_Level + 
                                      Marital_Status + 
                                      Income_Category + 
                                      Card_Category + 
                                      Months_on_book + 
                                      Total_Relationship_Count + 
                                      Months_Inactive_12_mon + 
                                      Contacts_Count_12_mon + 
                                      Credit_Limit + 
                                      Total_Revolving_Bal + 
                                      Avg_Open_To_Buy + 
                                      Total_Amt_Chng_Q4_Q1 + 
                                      Total_Trans_Amt + 
                                      Total_Trans_Ct + 
                                      Total_Ct_Chng_Q4_Q1 + 
                                      Avg_Utilization_Ratio, 
                                    data = train_data3)

customer_attrition_recipe4 <- customer_attrition_recipe4 %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_scale(all_numeric(), -all_outcomes()) %>%  
  step_center(all_numeric(), -all_outcomes())%>%
  step_impute_median(all_numeric(), -all_outcomes())

```


```{r}
train_data$Attrition_Flag <- as.factor(train_data$Attrition_Flag)
prepared_recipe4 <- prep(customer_attrition_recipe4, training = train_data3)

processed_train_data4 <- bake(prepared_recipe4, new_data = train_data)

```

```{r}
cv_set4 <- vfold_cv(train_data3, v = 10, strata = Attrition_Flag)

knn_model <- nearest_neighbor(neighbors = tune()) %>%
  set_engine("kknn") %>% 
  set_mode("classification")

knn_workflow <- workflow() %>%
  add_recipe(customer_attrition_recipe4) %>%
  add_model(knn_model)

knn_grid <- tibble(neighbors = seq(3, 15, by = 2))

knn_res <- knn_workflow %>%
  tune_grid(
    resamples = cv_set4,
    grid = knn_grid,
    control = control_grid(save_pred = TRUE),
    metrics = metric_set(roc_auc))

```

```{r}
best_knn <- knn_res %>%
  select_best(metric = "roc_auc")

final_knn_workflow <- knn_workflow %>%
  finalize_workflow(best_knn)

final_knn_fit <- final_knn_workflow %>%
  fit(data = train_data3)
```

```{r}
training_set$Attrition_Flag <- as.character(training_set$Attrition_Flag)
prepared_recipe3 <- prep(customer_attrition_recipe4, training = train_data3)
processed_train_data3 <- bake(prepared_recipe3, new_data = train_data3)

test_baked <- bake(prepared_recipe3, new_data = test_data2)

knn_test_pred <- predict(final_knn_fit, new_data = test_data2, type = "prob")

test_results_knn <- test_data2 %>%
  dplyr::select(Attrition_Flag) %>%
  bind_cols(knn_test_pred)

test_results_knn <- test_results_knn %>%
  rename(pred_Attrited_Customer = `.pred_Attrited Customer`)

test_results_knn$pred_class <- ifelse(test_results_knn$pred_Attrited_Customer > 0.5, 
                                      "Attrited Customer", 
                                      "Existing Customer")
```

```{r}

test_results_knn$pred_class <- factor(test_results_knn$pred_class, levels = c("Existing Customer", "Attrited Customer"))
test_results_knn$Attrition_Flag <- factor(test_results_knn$Attrition_Flag, levels = c("Existing Customer", "Attrited Customer"))
test_results_knn$pred_class <- factor(test_results_knn$pred_class, levels = c("Existing Customer", "Attrited Customer"))

accuracy_knn <- accuracy_vec(truth = test_results_knn$Attrition_Flag, estimate = test_results_knn$pred_class)


```

```{r}
roc_auc_knn <- roc_auc_vec(truth = test_results_knn$Attrition_Flag, 
  estimate = test_results_knn$pred_Attrited_Customer, event_level = "second")

print(paste("ROC AUC:", roc_auc_knn))

print(paste("Accuracy:", accuracy_knn))

```

```{r}
roc_curve_knn <- roc_curve(
  test_results_knn, 
  truth = Attrition_Flag, 
  pred_Attrited_Customer,
  event_level = "second" 
)
roc_auc_knn <- roc_auc(test_results_knn, truth = Attrition_Flag, pred_Attrited_Customer)

autoplot(roc_curve_knn) + 
  ggtitle(paste("ROC Curve for KNN Model", sep = "")) + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 
```
